---
title: "README"
author: "YZ"
date: "September 20, 2019"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting and Cleaning Data Course Project

set wd

```{r }
setwd("C:/Users/Public/Dropbox (Personal)/ETELSE/coursera/data science specialization/ds")
```

downloading and unzipping the dataset

```{r}
if(!file.exists(".")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl,destfile="./Dataset.zip")
```

Unzip dataSet to /data directory

```{r}
unzip(zipfile="./Dataset.zip",exdir="./data")
```

Reading trainings tables

```{r}
x_train <- read.table("./data/UCI HAR Dataset/train/X_train.txt")
y_train <- read.table("./data/UCI HAR Dataset/train/y_train.txt")
subject_train <- read.table("./data/UCI HAR Dataset/train/subject_train.txt")
```

Reading testing tables

```{r}
x_test <- read.table("./data/UCI HAR Dataset/test/X_test.txt")
y_test <- read.table("./data/UCI HAR Dataset/test/y_test.txt")
subject_test <- read.table("./data/UCI HAR Dataset/test/subject_test.txt")
```

Reading feature vector

```{r}
features <- read.table('./data/UCI HAR Dataset/features.txt')
```

Reading activity labels

```{r}
activityLabels = read.table('./data/UCI HAR Dataset/activity_labels.txt')
```

assigning column names

```{r}
colnames(x_train) <- features[,2] 
colnames(y_train) <-"activityId"
colnames(subject_train) <- "subjectId"

colnames(x_test) <- features[,2] 
colnames(y_test) <- "activityId"
colnames(subject_test) <- "subjectId"

colnames(activityLabels) <- c('activityId','activityType')
```

### 1. Merging the training and the test sets to create one data set

```{r}
mrg_train <- cbind(y_train, subject_train, x_train)
mrg_test <- cbind(y_test, subject_test, x_test)
setAllInOne <- rbind(mrg_train, mrg_test)
```

Reading column names

```{r}
colNames <- colnames(setAllInOne)
```

Create vector for defining ID, mean and standard deviation

```{r}
mean_and_std <- (grepl("activityId" , colNames) | 
                   grepl("subjectId" , colNames) | 
                   grepl("mean.." , colNames) | 
                   grepl("std.." , colNames) 
)
```

### 2. Extracting only the measurements on the mean and standard deviation for each measurement

```{r}
setForMeanAndStd <- setAllInOne[ , mean_and_std == TRUE]
```

### 3. Using descriptive activity names to name the activities in the data set

```{r}
setWithActivityNames <- merge(setForMeanAndStd, activityLabels,
                              by='activityId',
                              all.x=TRUE)
```
### 4. Appropriately labels the data set with descriptive variable names

This is done as part of the data prep for q.3

### 5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r}
TidySet <- aggregate(. ~subjectId + activityId, setWithActivityNames, mean)
TidySet <- TidySet[order(TidySet$subjectId, TidySet$activityId),]
write.table(TidySet, "TidySet.txt", row.name=FALSE)
```

### code book:

The new variables created are described above in preparation for q.3. The  A full description of the data is available at the site where the data was obtained: 

http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones